Making Sequential Programs Deterministic

### 2019-04-05 introduction.tex
Edits in sections 2.1 and 2.2
intro edits

We have created an analysis that detects nondeterminism or verifies its
absence.
Our analysis permits a programmer to specify which parts of their program
are intentionally nondeterministic, and it verifies that the remainder is deterministic.
%
If our analysis issues no warnings, then the program produces the same
output when executed twice on the same inputs.  This guarantee is modulo
the limitations of the
analysis, notably concurrency, implicit control flow, and unanalyzed libraries (see \cref{sec:threats}).
%
Our analysis works at compile time, giving a guarantee over every possible
execution of the program, by contrast to unsound dynamic tools that attempt
to discover when a program has exhibited nondeterministic behavior on
specific runs.  There is no need for a special JVM nor rerunning a program
multiple times.
%
Our analysis handles collections that will contain the same values, but
possibly in a different order, on different runs.
%
Our analysis is precise enough for practical use.  It permits calls to
nondeterministic APIs, and only issues a warning if they are used in ways
that may lead to nondeterministic output observed by a user.  Like any
sound analysis, it can issue false positive warnings.

The high-level goal of our work is to provide programmers with a tool for
specifying deterministic properties in a program and verifying them
statically.
%
Other researchers have also recognized the importance of the problem of nondeterminism.
Previous work in program analysis for nondeterminism has focused on unsound dynamic
approaches that identify flaky test cases.
NonDex uses a modified JVM that returns different results on different
executions, for a few key JDK methods with loose specifications.  Running a
test suite multiple times can reveal unwarranted dependence on those
methods.
%NonDex~\cite{nondex} is a tool that manually\todo{This is confusing:  the
%  tool can't manually identify.  Do you mean that the developers manually
%  identified, then the tool uses those models?}
%identifies 
DeFlaker~\cite{deflaker} looks at a range of commit versions
of a code, and marks a test as flaky if it doesn't execute any modified code but still fails in the newer version. These techniques
have been able to identify issues in real-world programs, some of which
have been fixed by the developers. We believe that identifying and
resolving nondeterminism
at an earlier phase in the software development lifecycle would be greatly beneficial to
developers.

@@

  \item We implemented analysis, as a pluggable type system for Java, in a
    tool called \theDeterminismChecker.

@@

  \item We compared \theDeterminismChecker our tool against state-of-the-art flaky test
    detectors, on their benchmarks.
    \TheDeterminismChecker found all the non-concurrency nondeterminism
    found by the other tools.

@@

\todo{It is essential that the introduction includes an example real-world
  defect that \theDeterminismChecker found.}

>>>>

We have created an analysis that detects nondeterminism or verifies its
absence in sequential programs.
Our analysis permits a programmer to specify which parts of their program
are intentionally nondeterministic, and it verifies that the remainder is deterministic.
%
If our analysis issues no warnings, then the program produces the same
output when executed twice on the same inputs.  This guarantee is modulo
the limitations of the
analysis, notably concurrency, implicit control flow, and unanalyzed libraries (see \cref{sec:threats}).
%
Our analysis works at compile time, giving a guarantee over every possible
execution of the program, by contrast to unsound dynamic tools that attempt
to discover when a program has exhibited nondeterministic behavior on
specific runs.  There is no need for a special JVM nor rerunning a program
multiple times.
%
Our analysis handles collections that will contain the same values, but
possibly in a different order, on different runs.
%
Our analysis is precise enough for practical use.  It permits calls to
nondeterministic APIs, and only issues a warning if they are used in ways
that may lead to nondeterministic output observed by a user.  Like any
sound analysis, it can issue false positive warnings.

The high-level goal of our work is to provide programmers with a tool for
specifying deterministic properties in a program and verifying them
statically.
%
Other researchers have also recognized the importance of the problem of nondeterminism.
Previous work in program analysis for nondeterminism has focused on unsound dynamic
approaches that identify flaky test cases.
NonDex~\cite{nondex} uses a modified JVM that returns different results on different
executions, for a few key JDK methods with loose specifications.  Running a
test suite multiple times can reveal unwarranted dependence on those
methods.
%NonDex~\cite{nondex} is a tool that manually\todo{This is confusing:  the
%  tool can't manually identify.  Do you mean that the developers manually
%  identified, then the tool uses those models?}
%identifies 
DeFlaker~\cite{deflaker} looks at a range of commit versions
of a code, and marks a test as flaky if it doesn't execute any modified code but still fails in the newer version. These techniques
have been able to identify issues in real-world programs, some of which
have been fixed by the developers. We believe that identifying and
resolving nondeterminism
at an earlier phase in the software development lifecycle would be greatly beneficial to
developers.

@@

  \item We implemented the analysis, as a pluggable type system for Java, in a
    tool called \theDeterminismChecker.

@@

  \item We compared our tool, \theDeterminismChecker, against state-of-the-art flaky test
    detectors, on their benchmarks.
    \TheDeterminismChecker found all the non-concurrency nondeterminism
    found by the other tools.

@@

\todo{1211 LOC of Java in the determinism checker}

%\todo{It is essential that the introduction includes an example real-world
%  defect that \theDeterminismChecker found.}
